Assessment and optimization of respiratory syncytial virus prophylaxis in Connecticut, \Sexpr{formatYear(startDate)}-\Sexpr{formatYear(endDate)}

### 2021-03-30 RSVOptimization.rnw
Separate main and supplement PDF
No abstract in supplement
TItle page for supplement
Reviewer feedback

\author[1,*]{Ben Artin, MEng, MPH, MMSc}
\author[1]{Virginia E. Pitzer, ScD}
\author[1]{Daniel M. Weinberger, PhD}
\affil[1]{Department of Epidemiology of Microbial Diseases, Yale School of Public Health, Yale University}
\affil[*]{ben@artins.org}

\begin{abstract}

\abstractsection{Background and Objectives}{Respiratory syncytial virus (RSV) causes seasonal respiratory infection, with hospitalization rates of up to 50\% in high-risk infants. Palivizumab provides safe and effective, yet costly, immunoprophylaxis. The American Academy of Pediatrics (AAP) recommends palivizumab only for high-risk infants and only during the RSV season. Outside of Florida, the current guidelines do not recommend regional adjustments to the timing of the immunoprophylaxis regimen. Our hypothesis is that adjusting the RSV prophylaxis regimen in Connecticut based on spatial variation in the timing of RSV incidence can reduce the disease burden compared to the current AAP-recommended prophylaxis regimen.}
\abstractsection{Methods}{We obtained weekly RSV-associated hospital admissions by ZIP-code in Connecticut between \Sexpr{formatMonth(startDate)} and \Sexpr{formatMonth(endDate)}. We estimated the fraction of all Connecticut RSV cases occurring during the period of protection offered by immunoprophylaxis (``preventable fraction'') under the AAP guidelines. We then used the same model to estimate protection conferred by immunoprophylaxis regimens with alternate start dates, but unchanged duration.}
\abstractsection{Results}{The fraction of RSV hospitalizations preventable by the AAP guidelines varies by county because of variations in epidemic timing. Prophylaxis regimens adjusted for state- or county-level variation in the timing of RSV seasons are modestly superior to the AAP-recommended regimen. The best alternative strategy yielded a preventable fraction of \Sexpr{regionStrategyCoverage("all", "stateMiddle", 1)} (\Sexpr{regionStrategyCoverageCI("all", "stateMiddle", 1, level)}), compared to \Sexpr{regionStrategyCoverage("all", "aap", 0)} (\Sexpr{regionStrategyCoverageCI("all", "aap", 0, level)}) for the AAP recommendation.}
\abstractsection{Conclusions}{Initiating RSV prophylaxis based on state-level data may improve protection compared with the AAP recommendations. In Connecticut, county-level recommendations would provide only a modest additional benefit while adding complexity.}

\end{abstract}

@@

\clearpage

\section{Introduction}

RSV causes upper respiratory tract infections in people of all ages. In adults, RSV infection is typically asymptomatic or gives rise to mild respiratory illness. However, RSV infection often incites more serious respiratory illness in children, with hospitalization rates as high as 4.4\% in infants with no comorbidities, making RSV a leading cause of hospitalization among infants.\cite{Hall:2009gi,Hall:2013jo,Zhou:2012ho,Boyce:2000gz}

@@

@@

In summary, our analysis reveals the potential for a $\sim 1\%$ gain in preventable RSV cases in Connecticut when a locally-tailored two-week (rather than monthly) prophylaxis guideline is used in place of the established AAP guidelines. At an average of $\Sexpr{prettyNum(totalRSV / totalYears, digits=3)}$ RSV-associated hospitalizations per year in children under 2 years old in Connecticut, this corresponds to $\sim \Sexpr{prettyNum(totalRSV / totalYears / 100, digits=3)}$ hospitalizations prevented per year, and an unknown number of prevented cases not necessitating hospitalization. Practical aspects of this potential gain --- such as implementation cost --- require further investigation. Additionally, the method we developed for this analysis can be applied to county-level or state-level analysis of seasonal RSV patterns in other regions, some of which are likely to show a greater relative benefit (due to a greater degree of misalignment between the RSV season and the AAP guidelines' prophylaxis window) as well as a greater absolute benefit (due to population size) than Connecticut.

@@

@@

\end{document}

>>>>

\author[1,*]{Ben Artin, MEng, MPH, MMSc}
\author[1]{Virginia E. Pitzer, ScD}
\author[1]{Daniel M. Weinberger, PhD}
\affil[1]{Department of Epidemiology of Microbial Diseases, Yale School of Public Health, Yale University}
\affil[*]{Correspondence to ben@artins.org}

% No abstract in supplement
\ifdef\skipmain{\def\noabstract}{
\begin{abstract}
\abstractsection{Background and Objectives}{Respiratory syncytial virus (RSV) causes seasonal respiratory infection, with hospitalization rates of up to 50\% in high-risk infants. Palivizumab provides safe and effective, yet costly, immunoprophylaxis. The American Academy of Pediatrics (AAP) recommends palivizumab only for high-risk infants and only during the RSV season. Outside of Florida, the current guidelines do not recommend regional adjustments to the timing of the immunoprophylaxis regimen. Our hypothesis is that adjusting the RSV prophylaxis regimen in Connecticut based on spatial variation in the timing of RSV incidence can reduce the disease burden compared to the current AAP-recommended prophylaxis regimen.}
\abstractsection{Methods}{We obtained weekly RSV-associated hospital admissions by ZIP-code in Connecticut between \Sexpr{formatMonth(startDate)} and \Sexpr{formatMonth(endDate)}. We estimated the fraction of all Connecticut RSV cases occurring during the period of protection offered by immunoprophylaxis (``preventable fraction'') under the AAP guidelines. We then used the same model to estimate protection conferred by immunoprophylaxis regimens with alternate start dates, but unchanged duration.}
\abstractsection{Results}{The fraction of RSV hospitalizations preventable by the AAP guidelines varies by county because of variations in epidemic timing. Prophylaxis regimens adjusted for state- or county-level variation in the timing of RSV seasons \cbstart are superior \cbend to the AAP-recommended regimen. The best alternative strategy yielded a preventable fraction of \Sexpr{regionStrategyCoverage("all", "stateMiddle", 1)} (\Sexpr{regionStrategyCoverageCI("all", "stateMiddle", 1, level)}), compared to \Sexpr{regionStrategyCoverage("all", "aap", 0)} (\Sexpr{regionStrategyCoverageCI("all", "aap", 0, level)}) for the AAP recommendation.}
\abstractsection{Conclusions}{\cbstart In Connecticut, county-level recommendations would provide only a minimal additional benefit while adding complexity. Initiating RSV prophylaxis based on state-level data may improve protection compared with the AAP recommendations. \cbend}

\end{abstract}
}

@@

\ifdef\skipmain{}{
\clearpage
\section{Introduction}

RSV causes upper respiratory tract infections in people of all ages. In adults, RSV infection is typically asymptomatic or gives rise to mild respiratory illness. However, RSV infection often incites more serious respiratory illness in children, with hospitalization rates as high as 4.4\% in infants with no comorbidities, making RSV a leading cause of hospitalization among infants.\cite{Hall:2009gi,Hall:2013jo,Zhou:2012ho,Boyce:2000gz} \cbstart Adults act as a reservoir of RSV and thereore play an important role in RSV mortality and morbidity in infants. \cbend

@@

\cbstart Further improvement in disease burden could likely be attained by extending the duration of the prophylaxis regimen to cover both the start and the end of the RSV season in the most populous counties (especially Fairfield county), but doing so would likely require extending the prophylaxis regimen by one additional dose and therefore increae the overall cost. \cbend

@@

In summary, our analysis reveals the potential for a $\sim 1\%$ gain in preventable RSV cases in Connecticut when a locally-tailored two-week (rather than monthly) prophylaxis guideline is used in place of the established AAP guidelines. At an average of $\Sexpr{prettyNum(totalRSV / totalYears, digits=3)}$ RSV-associated hospitalizations per year in children under 2 years old in Connecticut, this corresponds to $\sim \Sexpr{prettyNum(totalRSV / totalYears / 100, digits=3)}$ hospitalizations prevented per year, and an unknown number of prevented cases not necessitating hospitalization. Practical aspects of this potential gain --- such as implementation cost --- require further investigation.

\cbstart Additionally, the method we developed for this analysis can be applied to county-level or state-level analysis of seasonal RSV patterns in other regions, some of which are likely to show a greater relative benefit (due to a greater degree of misalignment between the RSV season and the AAP guidelines' prophylaxis window) as well as a greater absolute benefit (due to population size) than Connecticut. In particular, states of the US Pacific Northwest would likely benefit more than Connecticut, due to their later RSV season.\cite{Weinberger:2015hj} A similar approach is applicable to any other intervention against RSV that is sensitive to timing, including maternal vaccines and novel chemoprophylaxis. \cbend

@@

} % main

\ifdef\skipsupplement{}{

@@

} % supplement
\end{document}
